A Survey of Range-Only SLAM Techniques and Sensors

### 2018-08-20 Paper.tex
Added comments for future changes
Added some comments on how to make future changes and revised some grammer/structure

	\begin{abstract}		
		A common aspect in many autonomous robotics applications is addressing the problem of Simultaneous Localization and Mapping (SLAM). In most SLAM applications, the robot identifies distinguishable features in the environment, known as landmarks, within the environment through sensors such as cameras and laser rangefinders. Sensing these landmarks provide the robot with information such as range and bearing, which are used in the estimation proses of the robot's pose and the landmark locations within the map. However, some applications use landmarks which are only capable of providing range information, such as wireless and acoutstic beacons. In this survey, the variety of techniques used to solve the SLAM problem with respect to range only sensors (RO-SLAM) is explored.
	\end{abstract}

	\section{Introduction} 
		The Simultaneous Localization and Mapping (SLAM) problem is a common aspect of many autonomous robotics applications. SLAM involves estimating the pose of a mobile robot and building a map of its environment with no a priori information. The SLAM problem is a chicken-and-egg type problem, where the pose of the robot has to be with respect to a global frame, or map, and building a map requires knowledge of the robot's pose. There are many algorithms used to solve this problem, most using a landmark-based approach. Landmarks are distinguishable feature in the environment which can be observed through devices such as a camera or a laser rangefinder. Said devices are used to estimate the distance and orientation of the landmark relative to the robot. As the robot traverses the environment it continues to observe and collect landmarks. The sensed distance and angle from each landmark is combined with odometry readings to get an refined estimation of the robot pose and map. 
		%It can continue observing that landmark, as well as collecting new landmarks, and using the sensed distances and angles combined with odometry readings to refine its estimation of the robot's pose and map. 
		In most SLAM applications, references to the map is synonymous with the locations and features of the landmarks.

		In outdoor SLAM applications, such as self-driving cars and autonomous farming equipment, global positioning satellite (GPS) readings are used in addition to the vehicle's odometry. While GPS has low resolution, it proves useful as it provides location information, which can help compensate for the accumulated error from drift in odometric sensors. However, GPS isn't available in all environments, such as indoors and underwater. Instead, wireless signals and sonar beacons are often used in indoors and underwater environments, respectively. These beacons provide distance estimations to the robotic vehicle, but do not inherently provide location information as with GPS. However, these distances are often still useful as they can be used to compensate for the drift that occurs from measuring a vehicle's displacement through integration of inertial sensor readings. Most SLAM algorithms utilize these range values as a corrective measurement, the same way traditional algorithms use sensed landmarks. However, rather than using range and orientation to landmarks, beacons are only capable of providing range. SLAM algorithms which utilize these beacons are referred to as Range-Only SLAM (RO-SLAM).

		While the loss of orientation from traditional, vision based landmarks prevents refinement of the robot's orientation, range only beacons solve two major problems encountered in traditional landmark-based approaches: landmark association and landmark scaling. The landmark association problem occurs when a detected landmark has similar features to previously recorded landmark and is mistaken for the other, generating huge errors in the estimation of both the robot and the landmark. This problem is avoided using beacons, as they have some kind of network address which is broadcast during transmissions, so there is no way to confuse which range is associated with which landmark. The scaling problem occurs when a SLAM algorithm runs for a long time and more and more landmarks are appended to the map. As more landmarks are added, the amount of memory needed to store the positions and distributions of these landmarks increases, as well as computation time. When using beacons, a finite number of landmarks exists, and the total size of the map is known before deployment. In addition, fewer landmarks are needed in beacon-based applications, as most beacons do not require line-of-sight to be measured, unlike visual landmarks. Since the robot can observe the landmarks more often, it doesn't need to add nearly as many to ensure accurate position estimation.

		This paper surveys state of the art RO-SLAM applications. Before discussing RO-SLAM techniques, the different types of range sensors are introduced in Section \ref{range only sensors}, followed by initialization of range beacons in Section \ref{beacon initialization}. Solutions to the RO-SLAM state estimation problem are covered in Section \ref{methods and their implementations}. Finally, the conclusion in Section \ref{conclusion} closes with final thoughts regarding the state of the field of RO-SLAM.

	\section{range only sensors}
	\label{range only sensors}
		There are a variety of sensors which can be utilized as range beacons for RO-SLAM. These range sensors refer to any device which involves a signal transmission that can be used for the range estimation without line of sight. The most popular sensors can largely be broken down into two categories: acoustic based and radio frequency based. These two types of sensors both rely on the same principles for measuring distances, which is either by signal time of flight, or signal attenuation through propagation through each respective medium. These beacons serve as landmarks within larger SLAM algorithms. The following section surveys the different types of sensors which can be utilized as range beacons.

		\subsection{Acoustic}

			Acoustic ranging systems are primarily used with autonomous underwater vehicles (AUV) and sonar beacons that operate in underwater locations. The reason for this that radio waves experience strong attenuations, especially in salt water environments. As \cite{Partan2007} states, the issue of attenuations makes radio waves extremely impractical, thus making acoustics the logical alternative. Acoustic still suffer from a range of problems such as multipathing and propagation delays, but are one of the only meaningful alternatives. Papers such as those by \cite{Erol-Kantarci2011} and \cite{Akyildiz2005} explore a wide variety of issues and possible solutions that relate to acoustic sensors, such as the added complexity of assuming a 2D environment versus a 3D environment, or the design challenges that accompany underwater deployment with regards to spatial correlation and power consumption.

			A unique implementation of an underwater acoustic beacon is the Dive'N'Rise (DNR) beacon presented by \cite{Erol2007}. These DNR beacons rise to the surface, receive their location through GPS, then sink into the water and transmit their coordinates. The distance to an agent and the transmitting beacon can be determined via time of flight from the transmissions. A localization implementation using these beacons was accomplished by \cite{Erol2008}.

	%		  	\begin{figure}[h!]
	%				\centering
				%	\includegraphics[width=90mm]{DNR.png}
	%				\includegraphics[width=90mm]{DNR.png}	
	%				\caption{DNR being used to implement Dive and Rise Localization  \cite{Erol-Kantarci2011}}
	%				\label{DNR}
	%			\end{figure}

@@

			Radio frequency encompass a extensive spectrum of devices capable of transmitting or receiving using radio frequencies (RF). There are two methods primarily utilized to determine distance from radio frequency. One is by examining the received signal strength indicator (RSSI) and the other is the by measuring the time of flight of the signal. A system presented by \cite{Padmanabhan2000} processes RF signal strength and utilizes nearest neighbors heuristics to achieve localization based on said signal strength.  \cite{Kurth2003a} describes a method for using very small RF tags as beacons throughout the environment.  As a mobile robot moves throughout an environment it emits a query to which any RF tags that are within the transmission range respond. The robot then estimates the distance to the tag based on the delay. The response from the RF tags include a data tag which contains a unique ID number.

>>>>

	%The end of the abstract, I think, needs a better explanation of the survey other than stating that RO-SLAM techniques will be explored, ie: a more unifying theme
	\begin{abstract}		
		A common aspect in many autonomous robotics applications is addressing the problem of Simultaneous Localization and Mapping (SLAM). In most SLAM applications, the robot identifies distinguishable features in the environment, known as landmarks, through sensors such as cameras and laser rangefinders. Sensing these landmarks provide the robot with information such as range and bearing, which are used in the estimation processes of the robot's pose and the landmark locations within the map. However, some applications use landmarks which are only capable of providing range information, such as wireless and acoustic beacons. In this survey, the variety of techniques used to solve the SLAM problem with respect to range only sensors is explored.
	\end{abstract}

	\section{Introduction} 
		%Possible way to breakup RO-SLAM methods other than by sensor type: indoor and out door

		The Simultaneous Localization and Mapping (SLAM) problem is a common aspect of many robotics applications. SLAM involves estimating the pose of a mobile robot and building a map of its environment with no priori knowledge of said environment. The SLAM problem is a chicken-and-egg type problem, where the pose of the robot is ascertained with respect to a global frame, or map, and the building of a map requires knowledge of the robot's pose. There are many algorithms used to solve this problem, many use a landmark-based approach. Landmarks are distinguishable feature in an environment which can be observed via devices such as a camera or laser rangefinder. Said devices can be used to estimate the distance and orientation of the landmark relative to the robot. As the robot traverses the environment it continues to observe and collect landmarks. The sensed distance and orientation from each landmark is combined with odometry readings to get an refined estimation of the robot pose and map. 
		%It can continue observing that landmark, as well as collecting new landmarks, and using the sensed distances and angles combined with odometry readings to refine its estimation of the robot's pose and map. 
		In most SLAM applications, references to the map is synonymous with the locations and features of the landmarks.

		%A bit of a jarring transition. The goal is to introduce RO-SLAM and explain the difference between RO-SLAM and SLAM. The paragraph acomplishes that but maybe there is a smoother way to go about it. Also might need an explanation of location information and specifically how it differentiates from distance information
		In outdoor SLAM applications, such as self-driving cars and autonomous farming equipment, global positioning satellite (GPS) readings are used in addition to the vehicle's odometry. %ADD A CITATION
		While GPS has low resolution, it proves useful as it provides location information, which can help compensate for the accumulated error from drift in odometric sensors. However, GPS isn't available in all environments, such as indoors and underwater. Instead, wireless signals and sonar beacons are often used in indoors and underwater environments, respectively. These beacons provide distance estimations to the robotic vehicle, but do not inherently provide location information as with GPS. However, these distances are often still useful as they can be used to compensate for the drift that occurs from measuring a vehicle's displacement through integration of inertial sensor readings. Most SLAM algorithms utilize these range values as a corrective measurement, the same way traditional algorithms use sensed landmarks. However, rather than using range and orientation to landmarks, beacons are only capable of providing range. SLAM algorithms which utilize these beacons are referred to as Range-Only SLAM (RO-SLAM).

		While the loss of orientation from traditional, vision based landmarks  proves to be a detriment, range only beacons solve two major problems encountered in traditional landmark-based approaches: landmark association and landmark scaling. The landmark association problem occurs when a detected landmark has similar features to previously recorded landmark and is mistaken for the other, generating huge errors in the estimation of both the robot and the landmark. This problem is avoided using beacons, as they have some kind of network address which is broadcast during transmissions, so there is no way to confuse which range is associated with which landmark. The scaling problem occurs when a SLAM algorithm runs for a long time and more and more landmarks are appended to the map. As more landmarks are added, the amount of memory needed to store the positions and distributions of these landmarks increases, as well as computation time. When using beacons, a finite number of landmarks exists, and the total size of the map is known before deployment. In addition, fewer landmarks are needed in beacon-based applications, as most beacons do not require line-of-sight to be measured, unlike visual landmarks. Since the robot can observe the landmarks more often, it doesn't need to add nearly as many to ensure accurate position estimation. %Need a citatation to back that claim up

		%Eh... probably some better phrasing for this paragraph
		This paper surveys state of the art RO-SLAM applications. Before discussing RO-SLAM techniques, the different types of range sensors are introduced in Section \ref{range only sensors}, followed by initialization of range beacons in Section \ref{beacon initialization}. Solutions to the RO-SLAM state estimation problem are covered in Section \ref{methods and their implementations}. Finally, the conclusion in Section \ref{conclusion} closes with final thoughts regarding the state of the field of RO-SLAM.

	\section{range only sensors}
	\label{range only sensors}
		There are a variety of sensors which can be utilized as range-only beacons for RO-SLAM. These range sensors refer to any device which involves a signal transmission that can be used for the range estimation without requiring line of sight. The most popular sensors can largely be broken down into two categories: acoustic based and radio frequency based. These two types of sensors both largely rely on the same principles for measuring distances, which are either signal time of flight, or signal attenuation through propagation through each respective medium. These beacons serve as the landmarks within the RO-SLAM algorithms and impact how the algorithms are implemented.

		%I think this section could use some more details.
		\subsection{Acoustic}

			Acoustic ranging systems are largely used with autonomous underwater vehicles (AUV) that operate in underwater locations. The reason for this is that radio waves experience strong attenuations in salt water environments. As Partan states, the issue of attenuation makes radio waves extremely impractical, thus making acoustics the logical alternative \cite{Partan2007}. Acoustics still suffer from a range of problems such as multipathing and propagation delays, but compared to the alternatives they perform adequately in underwater environments. Papers such as those by \cite{Erol-Kantarci2011} and \cite{Akyildiz2005} explore a wide variety of issues and possible solutions that relate to acoustic sensors, such as the added complexity of assuming a two dimensional versus a three dimensional environment, or the design challenges that accompany underwater deployment with regards to spatial correlation and power consumption.

		  	%Ehhhhh.. This section... I dunno if this section is necissary
			A unique implementation of an underwater acoustic beacon is the Dive'N'Rise (DNR) beacon presented by \cite{Erol2007}. These DNR beacons rise to the surface, receive their location through GPS, then sink into the water and transmit their coordinates. The distance between an agent and the transmitting beacon can be determined via time of flight from the transmissions and used in conjunction with the GPS location to determine a refined location for the beacon. A localization implementation using these beacons was accomplished by \cite{Erol2008}.

@@

			Radio frequency encompasses an extensive spectrum of devices capable of transmitting or receiving signals using radio frequencies (RF). There are two methods commonly utilized to determine distance from radio frequency. One is by examining the received signal strength indicator (RSSI) and the other is by measuring the time of flight of the signal. A system presented by \cite{Padmanabhan2000} processes RF signal strength and utilizes nearest neighbors heuristics to achieve localization based on said signal strength.  \cite{Kurth2003a} describes a method for using very small RF tags as beacons throughout the environment.  As a mobile robot moves throughout the environment it emits a query to which any RF tags that are within the transmission range respond. The robot then estimates the distance to the tag based on the delay. The response from the RF tags include a data tag which contains a unique ID number.
